
---
title: "Establishing Pipelines with targets in R"
author: "Group E"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---

<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
div.comment {background-color:#F0F6FF; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)

"#f3f0ff"
```

<br>

------------------------------------------------------------------------

## Introduction

This exercise will walk you through establishing reproducible workflows using the `targets` package in R.
You will learn:
1) How to create a pipeline to manage data processing efficiently
2) The key functions in the targets package
3) The importance of file and folder structure
We will use the built-in `iris` dataset to demonstrate these workflows.

<div class="comment">
### Why Use Pipelines?
Pipelines help automate workflows by caching intermediate steps and only rerunning parts of the analysis that have changed.
</div>

```{r}
# Install targets if not already installed
library(targets)
library(tidyverse)
library(ggplot2)

tar_option_set(packages = c("dplyr", "ggplot2"))
```

<br>

------------------------------------------------------------------------

## Defining a Simple Pipeline

We'll create a simple pipeline to read the `iris` dataset, filter it for one species (`setosa`), calculate summary statistics, and generate a plot.

<div class="comment">
### Pipeline Setup (`_targets.R`)
</div>

```{r}
use_targets()
```

<br>

------------------------------------------------------------------------

## Running the Pipeline

Once your pipeline is defined in `_targets.R`, you can run it using the following command:

```{r}
tar_make()
```

------------------------------------------------------------------------

## Viewing Results

You can access the results of any target using `tar_read()`.

```{r}
# Read the summary statistics
tar_read(summary)

# View the plot: What would you need to write inside this function to see the plot?
tar_read(##Insert here)

```

<br>

## Visualizing the Pipeline

You can visualize the dependency graph of your pipeline with `tar_visnetwork()`.

```{r}
tar_visnetwork()
```

------------------------------------------------------------------------

## Conclusion

In this tutorial, we created a simple reproducible workflow using the `targets` package in R. By using `targets`, you can automate and manage complex analyses, improving both efficiency and reproducibility.
